Properties of winning Iterated Prisoner's Dilemma strategies.

### 2020-03-30 paper/main.tex
Conclusion --> Discussion and a few more edits

Moreover, the manner in which as strategy achieves a given cooperation rate
relative to the tournament population average is important. Playing a strategy
that randomly cooperates with $C_mean$ is unlikely to be as effective as it will
randomly have cooperations matched with defections. In contrast, TFT naturally
achieves a cooperation rate near $C_mean$ by virtue of copying its opponent's
last move while also minimizing instances where it is exploited by an opponent
(cooperating while the opponent defects), at least in non-noisy tournaments.
TFT forces the opponent to pay back the (C, D) round with a (D, C) round before
returning to mutual cooperation. This explains why TFT performed well
in Axelrod's original tournaments as most strategies submitted to those tournaments
were typically cooperative and relatively few strategies used an easily exploitable
pattern. TFT does not appear in the top ranks in
these tournaments because it is too nice. Strategies like Grudger will always
defect after a fixed number of opponent defections, which allows them to
effectively exploit strategies like Alternator or stochastic strategies that
have a non-zero chance of cooperating after mutual defection, which TFT will
not do. Moreover in a noisy environment these strategies will naturally tend
toward always defecting, leading them to exploit strategies like Cooperator. In
such a scenario the noisy environment effectively voids Axelrod's rule to be nice,
allowing strategies to attempt exploitation, whereas in a noise-free environment,
exploitation is risky because several strategies exhibit a Grudger-like behavior,
reducing the overall value in attempting to exploit strategies like Cooperator.

@@

\section{Conclusion}\label{section:conclusion}

This manuscript has explored the performance of \numberofstrategies strategies
of the Iterated Prisoner's Dilemma in a large number of computer tournaments.
We analyzed and extracted the salient features of the best performing strategies
across various tournament types, casting the results in terms of Axelrod's
original suggested features of good IPD strategies. Moreover, our results
shed light on the historic performance of TFT, zero determinant strategies,
and memory-one strategies generally. Strategies need to match their play
to the cooperativeness of the tournament population and do so in a way that
prevents or minimizes exploitation. Overall we see that complex or clever
strategies can be effective, whether trained against a corpus of possible
opponents or purposely designed to mitigate the impact of noise such as
in the strategy DBS. Further, we showed that while the type of exploitation
attempted by ZD strategies is not typically effective in tournaments, more
sophisticated strategies capable of selectively exploiting weaker opponents
while mutually cooperating with stronger opponents can be highly successful.
This fact was also indicated numerically by the importance of the strategy
feature SSE in the analysis of strategy features,

Highly noisy or tournaments with short matches favor less cooperative strategies.
These environments mitigate the value of being nice. The
added uncertainty reduces the ability of maintaining mutual cooperation and
also triggers grudging strategies. Accordingly we find that in noisy tournaments
that the best performing players cooperate a lower rate than the tournament
population on average. Nevertheless we found some strategies designed or trained
for noisy environments were also highly ranked in noise-free tournaments. This
indicates that strategy complexity is not necessarily a liability, rather it
can confer adaptability to a more diverse set of environments.

>>>>

Moreover, the manner in which as strategy achieves a given cooperation rate
relative to the tournament population average is important. Playing a strategy
that randomly cooperates with $C_mean$ is unlikely to be as effective as it will
randomly have cooperations matched with defections. In contrast, TFT naturally
achieves a cooperation rate near $C_mean$ by virtue of copying its opponent's
last move while also minimizing instances where it is exploited by an opponent
(cooperating while the opponent defects), at least in non-noisy tournaments.
\footnote{This also explains why Tit For N Tats does not fare well -- it fails
to achieve the proper cooperation ratio.}
TFT forces the opponent to pay back the (C, D) round with a (D, C) round before
returning to mutual cooperation. This explains why TFT performed well
in Axelrod's original tournaments as most strategies submitted to those tournaments
were typically cooperative and relatively few strategies used an easily exploitable
pattern. TFT does not appear in the top ranks in
these tournaments because it is too nice. Strategies like Grudger will always
defect after a fixed number of opponent defections, which allows them to
effectively exploit strategies like Alternator or stochastic strategies that
have a non-zero chance of cooperating after mutual defection, which TFT will
not do. Moreover in a noisy environment these strategies will naturally tend
toward always defecting, leading them to exploit strategies like Cooperator. In
such a scenario the noisy environment effectively voids Axelrod's rule to be nice,
allowing strategies to attempt exploitation, whereas in a noise-free environment,
exploitation is risky because several strategies exhibit a Grudger-like behavior,
reducing the overall value in attempting to exploit strategies like Cooperator.

@@

\section{Discussion}\label{section:conclusion}

This manuscript has explored the performance of \numberofstrategies strategies
of the Iterated Prisoner's Dilemma in a large number of computer tournaments.
We analyzed and extracted the salient features of the best performing strategies
across various tournament types, casting the results in terms of Axelrod's
original suggested features of good IPD strategies. Moreover, our results
shed light on the historic performance of TFT, zero determinant strategies,
and memory-one strategies generally. Strategies need to match their play
to the cooperativeness of the tournament population and do so in a way that
prevents or minimizes exploitation. Overall we see that complex or clever
strategies can be effective, whether trained against a corpus of possible
opponents or purposely designed to mitigate the impact of noise such as
in the strategy DBS. Further, we showed that while the type of exploitation
attempted by ZD strategies is not typically effective in tournaments, more
sophisticated strategies capable of selectively exploiting weaker opponents
while mutually cooperating with stronger opponents can be highly successful.
This fact was also indicated numerically by the importance of the strategy
feature SSE in the analysis of strategy features. These results highlight
a central idea in evolutionary game theory in this context: the fitness
landscape is a function of the population (where fitness in this case is
tournament performance). While that may seem obvious now, it shows why
historical tournament results on small or arbitrary populations of strategies
have so often failed to produce generalizable results.

Highly noisy or tournaments with short matches favor less cooperative strategies.
These environments mitigate the value of being nice. Uncertainty enables
exploitation, reducing the ability of maintaining or enforcing mutual cooperation,
while triggering grudging strategies to switch from typically cooperating to
typically defecting. Accordingly we find that in noisy tournaments
that the best performing players cooperate a lower rate than the tournament
population on average. Nevertheless we found some strategies designed or trained
for noisy environments were also highly ranked in noise-free tournaments. This
indicates that strategy complexity is not necessarily a liability, rather it
can confer adaptability to a more diverse set of environments.
